On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
our approach -> begin

Multithreading in Prolog is the ability to concurrently perform
computations, in which each computation runs independently but shares
the program clauses. When multithreading is combined with tabling, we
have the best of both worlds, since we can exploit the combination of
higher procedural control with higher declarative semantics. To the
best of our knowledge, XSB~\cite{Marques-08} and Yap~\cite{Areias-12a}
were the only Prolog systems that were able to combine both
multithreading with tabling. Yap's threads have their own execution
stacks and only share the code area where predicates, records, flags
and other global non-backtrackable data are stored. For tabled
evaluation, a thread views its tables as private but, at the engine
level, Yap has three designs: (i) The {\bf No-Sharing (NS)} design,
where each thread allocates fully private tables for each new tabled
subgoal called during its computation. (ii) The {\bf Subgoal-Sharing
  (SS)}, where threads share part of the table space and the (iii) The
{\bf Full-Sharing (FS)}, where threads share the complete table space.

In this paper we discuss a novel approach for solving the problem of
supporting multithreaded batched scheduling in the FS design and we
present a performance analysis comparison between local scheduling
with batched scheduling. Experimental results show that despite the
extra data structures required to extend the FS design with batched
scheduling, the execution time of the FS design with batched
scheduling is still quite competitive when comparing against local
scheduling.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{No-Sharing, Subgoal-Sharing and Full-Sharing Designs}

@@

@@

Table~\ref{tab_batched_overhead} shows the overhead ratios, when
compared with the NS design with 1 thread (running with local
scheduling), for the NS, SS and FS designs (all running TabMalloc),
when running 1, 8, 16, 24 and 32 threads with local and batched
scheduling on the five sets of benchmarks. For each design, the table
has then two columns, a column with {\bf Local} that shows results for
the local scheduling and a column with {\bf Batched} with results for
batched scheduling. The overhead results presented in both {\bf Local}
and {\bf Batched} columns use as base time the execution times
presented in the {\bf NS} column of the Table~\ref{tab_benchs}.

\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling without TabMalloc) for the NS,
  SS, FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling on the five sets of
  benchmarks (best ratios by row and by design for the Minimum,
  Average and Maximum are in bold)}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{c}{\multirow{1}{*}{\bf SS}} & 
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-8}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& {\bf 0.54}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 0.84}& 0.90& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.04}& {\bf 1.04}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.17& 0.16& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 0.92}& 0.93& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& 1.20& {\bf 1.15}& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.15& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 0.82& {\bf 0.77}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.04}& 1.05& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& 1.31& {\bf 1.28}& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.12& 0.13& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.02& {\bf 0.98}& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& 1.22& {\bf 1.19}& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& 1.81& {\bf 1.61}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.18& 0.16& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& {\bf 1.07}& 1.12& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& 1.54& {\bf 1.51}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 2.52}& 2.62& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.42& 0.43& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\end{table}

By observing Table~\ref{tab_batched_overhead}, we can see that, for
one thread, on average, local scheduling is sightly better than
batched on the three designs. For the NS design we have $0.78$ and
$0.82$, for the SS design we have $0.84$ and $0.90$ and for the FS
design we have $1.30$ and $1.46$ average overhead ratios, for the
local and batched scheduling strategies, respectively. As we scale the
number of threads, one can observe that, for the NS and SS designs
both scheduling strategies have similar minimum, average and maximum
overhead ratios. For the FS design, the best minimum overhead ratio is
always for batched scheduling. For the average and maximum overhead
ratio, local scheduling is always better than batched scheduling. In
summary, we can say that both the local and batched scheduling
strategies have similar overhead results on worst case scenarios for
the NS, SS and FS designs.

>>>>

Multithreading in Prolog is the ability to concurrently perform
computations, in which each computation runs independently but shares
the program clauses. When multithreading is combined with tabling, we
have the best of both worlds, since we can exploit the combination of
higher procedural control with higher declarative semantics. To the
best of our knowledge, XSB~\cite{Marques-08} and Yap~\cite{Areias-12a}
were the only Prolog systems that were able to combine both
multithreading with tabling. Yap's threads have their own execution
stacks and only share the code area where predicates, records, flags
and other global non-backtrackable data are stored. For tabled
evaluation, a thread views its tables as private but, at the engine
level, Yap has three designs that vary from a \emph{No-Sharing} (NS)
design, where each thread allocates fully private tables for each new
tabled subgoal called during its computation, to a \emph{Full-Sharing}
(FS), where threads share the complete table space.

In this paper we discuss a novel approach for solving the problem of
supporting multithreaded batched scheduling in the FS design and we
present a performance analysis comparison between local scheduling
with batched scheduling. Experimental results show that despite the
extra data structures required to combine FS design with batched
scheduling, the execution time of combination is still quite
competitive when comparing against the FS with local scheduling.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Yap's No-Sharing and Full-Sharing Designs}

\begin{figure}[!ht]
\centering
\includegraphics[width=10.5cm]{figures/arch.pdf}
\caption{The \emph{FS} design with the PAC procedure - (a) private
  chaining and (b) public chaining}
\label{fig_tabtries_pcc}
\end{figure}

@@

Table~\ref{tab_batched_overhead} shows the overhead ratios, when
compared with the NS design with 1 thread (running with local
scheduling), for the NS and FS designs (all running TabMalloc), when
running 1, 8, 16, 24 and 32 threads with local and batched scheduling
on the five sets of benchmarks. For each design, the table has then
two columns, a column with {\bf Local} that shows results for the
local scheduling and a column with {\bf Batched} with results for
batched scheduling. The overhead results presented in both {\bf Local}
and {\bf Batched} columns use as base time the execution times
presented in the {\bf NS} column of the Table~\ref{tab_benchs}.

\setlength{\tabcolsep}{12pt}

\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling without TabMalloc) for the NS
  and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling on the five sets of
  benchmarks (best ratios by row and by design for the Minimum,
  Average and Maximum are in bold)}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\end{table}

By observing Table~\ref{tab_batched_overhead}, we can see that, for
one thread, on average, local scheduling is sightly better than
batched on the two designs. For the NS design we have $0.78$ and
$0.82$ and for the FS design we have $1.30$ and $1.46$ average
overhead ratios, for the local and batched scheduling strategies,
respectively. As we scale the number of threads, one can observe that,
for the NS design both scheduling strategies have similar minimum,
average and maximum overhead ratios. For the FS design, the best
minimum overhead ratio on $1$, $8$, $16$, $24$ and $32$ threads is
always the batched scheduling. For the average and maximum overhead
ratio, local scheduling is always better than batched
scheduling. However, for the average ratio, the results between both
strategies is quite close. We consider this to be quite significant,
once the FS design with batched scheduling requires the support of the
PAC procedure. For the maximum overhead ratio, the difference between
both strategies is higher than for the average overhead ratio, however
we hope to reduce the difference with further improvements in the PAC
procedure.
